A Horrocks-type theorem for even orthogonal $K_2$

### 2019-08-22 dl-paper.tex
Minor fixes

Thus, our main result is the following theorem, which is the orthogonal analogue of~\cite[Theorem~5.1]{Tu83} and the $\K_2$-analogue of~\cite[Theorem~6.8]{Su82} (cf. also with~\cite[Theorem~VI.5.2]{Lam10} and~\cite[Theorem~1.1]{St-poly}).
\begin{theorem}[Horrocks theorem for orthogonal $\K_2$]\label{thm:main} Let $A$ be a commutative ring in which $2$ is invertible. Then for any $\ell \geq 7$ the following commutative square is a pullback square in which all maps are injective:
\begin{equation} \nonumber
  \xymatrix{ \KO_2(2\ell, A) \ar[r] \ar[d] & \KO_2(2\ell, A[X]) \ar[d] \\ \KO_2(2\ell, A[X\inv]) \ar[r] & \KO_2(2\ell, A[X, X\inv]).}  
\end{equation} 
Moreover, the same assertion holds if one replaces the functor $\KO_2(2\ell, -)$ with $\K_2(\rD_\ell, -)$ or $\St(\rD_\ell, -)$. \end{theorem}
In the above statement $\KO_2(2\ell, -)$ denotes the unstable orthogonal $K_2$-functor (see~\cref{sec:quillen}).

@@

\subsection{Acknowledgements} \cref{thm41} is obtained by the first-named author who acknowledges the financial support of RSCF grant No ???.
\cref{thm:P1glueing} is obtained by the second-named author who acknowledges the financial support of JSC Gazpromneft and RFBR grant No 18-31-20044

@@

\subsection{Relative Steinberg groups and unstable K-groups} \label{sec:quillen}
In this subsection we recall the definitions and basic facts pertaining to the theory of relative extensions developed by J.-L.~Loday in~\cite{Lo78}. This theory was applied by J.~Loday to the stable Steinberg group $\St(R)$ from algebraic K-theory. The main goal of this subsection is to show that Loday's theory can be applied to unstable Steinberg groups, and that the resulting relative Steinberg groups enjoy the same good properties as their stable counterparts.
Some of the results of this subsection have been briefly mentioned in~\cite{S15} (cf. e.\,g. Corollaries 3--4).

@@

Notice that $\KO_2(2n, R) = \Ker(\St(\rD_\ell, R) \to \mathrm{SO}(2n, R))$ is a larger subgroup of $\St(\rD_\ell, R)$ than $\K_2(\rD_\ell, R)$.
More precisely, from~\cite[157]{Ba74} and Snake lemma it follows that \[\KO_2(2n, R)/\K_2(\rD_n, R) \cong \Ker(\mathrm{Spin}(2n, R) \to \mathrm{SO}(2n, R)) = \mu_2(R).\]
On the other hand, from~\cref{lem:lowerKgroups} it follows that $\KO_3(2n, R)$ is isomorphic to the group $K_3^{G, P}$ in which
$G = G(\rD_n, -) = \mathrm{Spin}(2n, -)$ as $P = \mathrm{Epin}(2n, -)$ is the elementary subfunctor of $G$.
%Notice that $\K_2(\rD_n, R) \subset \mathrm{KO}_2(n, R)$.

>>>>

Our main result is the following theorem, which is the orthogonal analogue of~\cite[Theorem~5.1]{Tu83} and the $\K_2$-analogue of~\cite[Theorem~6.8]{Su82} (cf. also with~\cite[Theorem~VI.5.2]{Lam10} and~\cite[Theorem~1.1]{St-poly}).
\begin{theorem}[Horrocks theorem for orthogonal $\K_2$]\label{thm:main} Let $A$ be a commutative ring in which $2$ is invertible. Then for any $\ell \geq 7$ the following commutative square is a pullback square in which all maps are injective:
\begin{equation} \nonumber
  \xymatrix{ \KO_2(2\ell, A) \ar[r] \ar[d] & \KO_2(2\ell, A[X]) \ar[d] \\ \KO_2(2\ell, A[X\inv]) \ar[r] & \KO_2(2\ell, A[X, X\inv]).}  
\end{equation} 
Moreover, the same assertion holds if one replaces the functor $\KO_2(2\ell, -)$ with $\K_2(\rD_\ell, -)$ or $\St(\rD_\ell, -)$. \end{theorem}
In the above statement $\KO_2(2\ell, -)$ denotes the unstable orthogonal $K_2$-functor (see~\cref{sec:quillen}).

@@

\subsection{Acknowledgements} \cref{thm41} was obtained by the first-named author who acknowledges the financial support of RSF grant No 17-11-01261.
\cref{thm:P1glueing} was obtained by the second-named author who acknowledges the financial support of JSC Gazpromneft and RFBR grant No 18-31-20044.

@@

\subsection{Relative Steinberg groups and unstable K-groups} \label{sec:quillen}
In this subsection we recall the definitions and basic facts pertaining to the theory of relative central extensions developed by J.-L.~Loday in~\cite{Lo78}. This theory was applied by J.~Loday to the stable Steinberg group $\St(R)$ from algebraic K-theory. The main goal of this subsection is to show that Loday's theory also can be applied to unstable Steinberg groups, and that the resulting relative Steinberg groups enjoy many good properties of their stable counterparts.
Some of the results of this subsection have been briefly mentioned in~\cite{S15} (cf. e.\,g. Corollaries 3--4).

@@

Notice that $\KO_2(2n, R) = \Ker(\St(\rD_\ell, R) \to \mathrm{SO}(2n, R))$ is a larger subgroup of $\St(\rD_\ell, R)$ than $\K_2(\rD_\ell, R)$.
More precisely, from~\cite[157]{Ba74} and Snake lemma it follows that \[\KO_2(2n, R)/\K_2(\rD_n, R) \cong \Ker(\mathrm{Spin}(2n, R) \to \mathrm{SO}(2n, R)) = \mu_2(R).\]
On the other hand, from~\cref{lem:lowerKgroups} it follows that $\KO_3(2n, R)$ is isomorphic to the group $K_3^{G, P}$ in which
$G = G(\rD_n, -) = \mathrm{Spin}(2n, -)$ as $P = \mathrm{Epin}(2n, -)$ is its elementary subfunctor.
%Notice that $\K_2(\rD_n, R) \subset \mathrm{KO}_2(n, R)$.
